Individual Report

### 2020-03-03 doc/RealEstate_Xuechun_Lu.Rmd
Add files via upload

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Summary
A real estate analyst in Vancouver was interested in predicting future property tax for his company by predicting future mill rates. His expectation was to obtain a good prediction model to predict mill rates in each municipality in Metro Vancouver given the past data. In this study, data cleaning was done to prepare for the data analysis and raw data was transformed into summary statistics for each municipality. Exploratory data analysis (EDA) shows that mill rate is strongly correlated with year, municipality, tax class, assessment total, improvement, land and  property count, and weakly correlated with municipal budget. Linear regression models (ordinary linear regression, Lasso, Ridge and Elastic Net regression) were built to explore the relationship between mill rate and potential features. Among all models, ordinary linear regression with reduced features gives the slightly better goodness of fit and prediction power.

@@

dat_pred_income <- assessment_aggregate %>% select(Year,AddressAssessorMunicipalityDesc,TaxClassCode,predTaxIncome)
dat_pred_income <- aggregate(dat_pred_income$predTaxIncome,by=list(Year=dat_pred_income$Year,Municipalities=dat_pred_income$AddressAssessorMunicipalityDesc),FUN=sum)
names(dat_pred_income)[3] <- paste("predTaxIncome")

@@

compare_full_plot <- income_budget_compare %>% filter(Municipalities!="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=Municipalities)) + geom_line()

compare_plot1 <- income_budget_compare %>% filter(Municipalities=="Vancouver") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s tax for Vancouver")
compare_plot2 <- income_budget_compare %>% filter(Municipalities=="Surrey") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s tax for Surrey")
compare_plot3 <- income_budget_compare %>% filter(Municipalities=="Burnaby") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s tax or Burnaby")
compare_plot4 <- income_budget_compare %>% filter(Municipalities=="Richmond") %>% ggplot(aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),color=TaxType)) + geom_line() + ggtitle("budget v.s tax for Richmond")
multiplot(compare_plot1,compare_plot2,compare_plot3,compare_plot4,cols=2)
```

@@

## Correlation for continuous variables
assess_class <- assessment_aggregate %>% ggplot(aes(x=assessTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and assess - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+theme(legend.position = "none")

tax_class <- assessment_aggregate %>% ggplot(aes(x=tax,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and tax - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+theme(legend.position = "none")

land_class<-assessment_aggregate %>% ggplot(aes(x=landTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and landTotal - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+theme(legend.position = "none")

improvementTotal_class<-assessment_aggregate %>% ggplot(aes(x=improvementTotal,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and improvementTotal - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+theme(legend.position = "none")

propertyCount_class<-assessment_aggregate %>% ggplot(aes(x=propertyCount,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate and propertyCount - cor") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+theme(legend.position = "none")

multiplot(assess_class,
tax_class,
land_class,improvementTotal_class,
          propertyCount_class, cols=2)

@@

Plot 2. scatterplot with regression line for mill rate against assessTotal, tax, landTotal, improvementTotal and propertyCount.

@@

# 2. Goodness of Fit
```{r, echo=T, results="hide"}
# full linear model
linear_full<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=assessment_aggregate)
summary(linear_full)
linear_full_fit<-augment(linear_full)
sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) 
# Multiple R-squared:  0.8874,	Adjusted R-squared:  0.8707 
# MSPE = 1.984285
row1<-c("OLR full",0.8874, 0.8707,round(1.984285,4))

# reduced model
reduced<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+assessTotal+landTotal, data=assessment_aggregate)
summary(reduced)
reduced_fit<-augment(reduced)
# Multiple R-squared:  0.8874,	Adjusted R-squared:  0.8721 
sqrt(sum((reduced_fit$.resid)^2)/nrow(assessment_aggregate))
row2<-c("OLR reduced",0.8874,0.8721,round(1.984541,4))
```
```{r, echo=F}
# create a list of significant variables from the full linear model

@@

```{r, echo=T, results="hide"}
# Ridge
dummy_year<-dummy(assessment_aggregate$Year)
dummy_municipal<-dummy(assessment_aggregate$AddressAssessorMunicipalityDesc)
dummy_taxclass<-dummy(assessment_aggregate$TaxClassCode)
# build x matrix
x<-cbind(dummy_municipal,dummy_year,dummy_taxclass,assessment_aggregate$assessTotal,assessment_aggregate$landTotal,assessment_aggregate$improvementTotal,assessment_aggregate$propertyCount,assessment_aggregate$tax)
y<-assessment_aggregate$rate
lambdas <- 10^seq(2, -3, by = -.1)

@@

# Elastic Net
#tibble::as_tibble(assessment_aggregate)
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
  rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount, data = assessment_aggregate,
  method = "glmnet",
  trControl = cv_10
)
elastic_net
row5<-c("Elastic Net",round(0.8625398,4),"N/A",round(2.236598,4))

@@

```{r, echo=T, results="hide", warning=F}
# full linear model
newx<-test[,-c(8,9)]
y<-test[,c(8)]
linear_1<-lm(rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax,data=train)
resid<-predict(linear_1,newdata = newx) - y
row1<-c("OLR full", round(sqrt(sum(resid^2)/nrow(test)),4))

# reduced model
linear_2<-lm(rate~factor(Year)+factor(TaxClassCode)+factor(AddressAssessorMunicipalityDesc)+assessTotal+landTotal,data=train)
resid<-predict(linear_2,newdata = newx) - y
row2<-c("OLR reduced",round(sqrt(sum(resid^2)/nrow(test)),4))
```

@@

# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x_train, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
lambda_best <- lasso_reg$lambda.min;lambda_best
lasso_model <- glmnet(x_train, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
eval_results(y_test, predictions_test)
row4<-c("Lasso",round(2.528047,4))

# ridge
ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
set.seed(450)
cv_ridge <- cv.glmnet(x_train, y_train, alpha = 0, lambda = lambdas, nfolds=10)
optimal_lambda <- cv_ridge$lambda.min
optimal_lambda
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
eval_results(y_test, predictions_test)
row3<-c("Ridge",round(2.567499,4))

# elastic net
tibble::as_tibble(assessment_aggregate[train_ind,])
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
 rate~factor(AddressAssessorMunicipalityDesc)+factor(Year)+factor(TaxClassCode)+assessTotal+landTotal+improvementTotal+propertyCount+tax, 
 data = assessment_aggregate[train_ind,],
  method = "glmnet",
  trControl = cv_10
)
elastic_net

@@

elastic_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 1, family = 'gaussian', lambda =  0.06549203)
predictions_test <- predict(elastic_reg, newx = x_test)
eval_results(y_test, predictions_test)
row5<-c("Elastic Net",round(2.54861,4))

@@

```{r, echo=F}
colsname<-c("Model", "MSPE")
table<-rbind(row1,row2,row3,row4,row5)
colnames(table)<-colsname
rownames(table)<-NULL
kable(table, caption="Prediction power for all models")
```

@@

The OLR shows significant variables related to mill rates are Year, Tax Class, Municipality, Assessment and Land Total

@@

@@

There are 1801 missing values in mill rate(TaxClassTaxRate). We decided to impute these missing values
Based on client information, all properties in the same region, classcode, and year should have a unique class rate

- For entries with mill rate, we aggregated them into groups by region + classcode + year.

- For entries without mill rate, we found the group they belong to and assign them mill rate in that group.

Here is some exceptions found:

Some groups' mill rate is not unique:

@@

>>>>

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```
# Summary
A real estate analyst in Vancouver was interested in predicting future property tax for his company by predicting future mill rates. His expectation was to obtain a good prediction model to predict mill rates in each municipality in Metro Vancouver given the past data. In this study, data cleaning was done to prepare for the data analysis and raw data was transformed into summary statistics for each municipality. Exploratory data analysis (EDA) shows that mill rate is strongly correlated with year, municipality, tax class, assessment total, improvement, land and  property count, and weakly correlated with municipal budget. Linear regression models (ordinary linear regression, Lasso, Ridge and Elastic Net regression) were built to explore the relationship between mill rate and potential features. Among all models, ordinary linear regression with reduced features gives the slightly better goodness of fit and prediction power.

@@

dat_pred_income <- assessment_aggregate %>% select(Year,
                                                   AddressAssessorMunicipalityDesc,
                                                   TaxClassCode,
                                                   predTaxIncome)
dat_pred_income <- aggregate(dat_pred_income$predTaxIncome,
                             by=list(Year=dat_pred_income$Year,
                                     Municipalities=dat_pred_income$AddressAssessorMunicipalityDesc),
                             FUN=sum)
names(dat_pred_income)[3] <- paste("predTaxIncome")

@@

compare_full_plot <- income_budget_compare %>% filter(Municipalities!="Vancouver") %>% 
  ggplot(
  aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),
      color=Municipalities)) + 
  geom_line()

compare_plot1 <- income_budget_compare %>% filter(Municipalities=="Vancouver") %>%
  ggplot(
    aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),
        color=TaxType)) + 
  geom_line() + 
  ggtitle("budget v.s tax for Vancouver")

compare_plot2 <- income_budget_compare %>% filter(Municipalities=="Surrey") %>% 
  ggplot(
  aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),
      color=TaxType)) + 
  geom_line() + 
  ggtitle("budget v.s tax for Surrey")

compare_plot3 <- income_budget_compare %>% filter(Municipalities=="Burnaby") %>% 
  ggplot(
  aes(x=Year,y=Amount,
      group=interaction(Municipalities,TaxType),
      color=TaxType)) +
  geom_line() + 
  ggtitle("budget v.s tax or Burnaby")

compare_plot4 <- income_budget_compare %>% filter(Municipalities=="Richmond") %>% 
  ggplot(
  aes(x=Year,y=Amount,group=interaction(Municipalities,TaxType),
      color=TaxType)) + 
  geom_line() + 
  ggtitle("budget v.s tax for Richmond")

multiplot(compare_plot1,compare_plot2,compare_plot3,compare_plot4,cols=2)
```

@@

## Correlation for continuous variables
assess_class <- assessment_aggregate %>% 
  ggplot(
  aes(x=assessTotal,
      y=rate,
      group=AddressAssessorMunicipalityDesc,
      color=AddressAssessorMunicipalityDesc)) + 
  geom_point() + 
  ggtitle("mill rate and assessTotal") + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+
  theme(legend.position = "none")

tax_class <- assessment_aggregate %>% 
  ggplot(
  aes(x=tax,
      y=rate,
      group=AddressAssessorMunicipalityDesc,
      color=AddressAssessorMunicipalityDesc)) + 
  geom_point() + 
  ggtitle("mill rate and budget") +
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+
  theme(legend.position = "none")

land_class<-assessment_aggregate %>% ggplot(
  aes(x=landTotal,
      y=rate,
      group=AddressAssessorMunicipalityDesc,
      color=AddressAssessorMunicipalityDesc)) + 
  geom_point() + 
  ggtitle("mill rate and landTotal") + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+
  theme(legend.position = "none")

improvementTotal_class<-assessment_aggregate %>% 
  ggplot(
  aes(x=improvementTotal,
      y=rate,
      group=AddressAssessorMunicipalityDesc,
      color=AddressAssessorMunicipalityDesc)) + 
  geom_point() + 
  ggtitle("mill rate and improvementTotal") + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+
  theme(legend.position = "none")

propertyCount_class<-assessment_aggregate %>% 
  ggplot(
  aes(x=propertyCount,
      y=rate,
      group=AddressAssessorMunicipalityDesc,
      color=AddressAssessorMunicipalityDesc)) + 
  geom_point() + 
  ggtitle("mill rate and propertyCount") + 
  geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")+
  theme(legend.position = "none")

multiplot(
  assess_class,
  tax_class,
  land_class,
  improvementTotal_class,
  propertyCount_class, 
  cols=2)

@@

Plot 2. scatterplot with regression line for mill rate against assessTotal, municial budget, landTotal, improvementTotal and propertyCount.

@@

# 2. Goodness of Fit
```{r, echo=T, results="hide"}
# full linear model
linear_full<-lm(rate~
                  factor(AddressAssessorMunicipalityDesc)+
                  factor(Year)+
                  factor(TaxClassCode)+
                  assessTotal+
                  landTotal+
                  improvementTotal+
                  propertyCount+
                  tax,
                data=assessment_aggregate)

summary(linear_full)
linear_full_fit<-augment(linear_full)
sqrt(sum((linear_full_fit$.resid)^2)/nrow(assessment_aggregate)) 
# Multiple R-squared:  0.8874,	Adjusted R-squared:  0.8707 
# MSPE = 1.984285
row1<-c("OLR full",0.8874, 0.8707,round(1.984285,4))

# reduced model
reduced<-lm(rate~
              factor(Year)+
              factor(TaxClassCode)+
              factor(AddressAssessorMunicipalityDesc)+
              assessTotal+
              landTotal, 
            data=assessment_aggregate)
summary(reduced)
reduced_fit<-augment(reduced)
# Multiple R-squared:  0.8874,	Adjusted R-squared:  0.8721 
sqrt(sum((reduced_fit$.resid)^2)/nrow(assessment_aggregate))
row2<-c("OLR reduced",0.8874,0.8721,round(1.984541,4))
```

```{r, echo=F}
# create a list of significant variables from the full linear model

@@

```{r, echo=T, results="hide"}
# Ridge
dummy_year<-dummy(assessment_aggregate$Year)
dummy_municipal<-dummy(assessment_aggregate$AddressAssessorMunicipalityDesc)
dummy_taxclass<-dummy(assessment_aggregate$TaxClassCode)
# build x matrix
x<-cbind(dummy_municipal,
         dummy_year,
         dummy_taxclass,
         assessment_aggregate$assessTotal,
         assessment_aggregate$landTotal,
         assessment_aggregate$improvementTotal,
         assessment_aggregate$propertyCount,
         assessment_aggregate$tax)

y<-assessment_aggregate$rate
lambdas <- 10^seq(2, -3, by = -.1)

@@

# Elastic Net
#tibble::as_tibble(assessment_aggregate)
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
  rate~
    factor(AddressAssessorMunicipalityDesc)+
    factor(Year)+
    factor(TaxClassCode)+
    assessTotal+
    landTotal+
    improvementTotal+
    propertyCount, 
  data = assessment_aggregate,
  method = "glmnet",
  trControl = cv_10
)
elastic_net
row5<-c("Elastic Net",round(0.8625398,4),"N/A",round(2.236598,4))

@@

```{r, echo=T, results="hide", warning=F}
# full linear model
newx<-test[,-c(8,9)]
y<-test[,c(8)]
linear_1<-lm(rate~
               factor(AddressAssessorMunicipalityDesc)+
               factor(Year)+
               factor(TaxClassCode)+
               assessTotal+
               landTotal+
               improvementTotal+
               propertyCount+
               tax,
             data=train)
resid<-predict(linear_1,newdata = newx) - y

# reduced model
linear_2<-lm(rate~
               factor(Year)+
               factor(TaxClassCode)+
               factor(AddressAssessorMunicipalityDesc)+
               assessTotal+
               landTotal,
             data=train)

resid<-predict(linear_2,newdata = newx) - y
```

@@

# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x_train, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)
lambda_best <- lasso_reg$lambda.min;lambda_best
lasso_model <- glmnet(x_train, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)
predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
eval_results(y_test, predictions_test)

# ridge
ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
set.seed(450)
cv_ridge <- cv.glmnet(x_train, y_train, alpha = 0, lambda = lambdas, nfolds=10)
optimal_lambda <- cv_ridge$lambda.min
optimal_lambda
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
eval_results(y_test, predictions_test)

# elastic net
tibble::as_tibble(assessment_aggregate[train_ind,])
cv_10 = trainControl(method = "cv", number = 10)
elastic_net = train(
 rate~
   factor(AddressAssessorMunicipalityDesc)+
   factor(Year)+
   factor(TaxClassCode)+
   assessTotal+landTotal+
   improvementTotal+
   propertyCount+
   tax, 
 data = assessment_aggregate[train_ind,],
  method = "glmnet",
  trControl = cv_10
)
elastic_net

@@

elastic_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 1, family = 'gaussian', lambda =  0.06549203)
predictions_test <- predict(elastic_reg, newx = x_test)
eval_results(y_test, predictions_test)

@@

```{r, echo=F}
colsname<-c("Model", "MSPE")
row1<-c("OLR full", round(sqrt(sum(resid^2)/nrow(test)),4))
row2<-c("OLR reduced",round(sqrt(sum(resid^2)/nrow(test)),4))
row3<-c("Ridge",round(2.567499,4))
row4<-c("Lasso",round(2.528047,4))
row5<-c("Elastic Net",round(2.54861,4))
table<-rbind(row1,row2,row3,row4,row5)
colnames(table)<-colsname
rownames(table)<-NULL
kable(table, caption="Prediction power for all models")
```

@@

The OLR shows significant variables related to mill rates are Year, Tax Class, Municipality, Assessment and Land Total.

@@

In the current design, prediction power is measured by MSPE on one trial of testing set, which might cause a large variance of MSPE and thus an inaccurate prediction power. In the next stage, in order to get a more accurate prediction power, we will perform k-fold cross validation to decrease the variance of MSPE.

@@

1. Mill Rate

There are 1801 missing values in mill rate(TaxClassTaxRate). Since properties belonging to the same group (same year, region and tax class) should share a common mill rate, missing values were imputed by the common mill rate of properties in the same group.

However, some groups' mill rates are not unique or missing entirely. The treatments are as below:

@@

2. Municipal Budget

We used total tax collected in each area to approximate municipal budget. However, the governement website only has data available up to 2018; total tax collected in 2019 is missing. In this study, tax collected in 2019 is imputed by average value of tax collected in the previous three years.
